An agent-based model of local food systems

### 2016-05-13 draft.tex
update

\subsubsection{Others $(p, c, \alpha, \beta, \gamma, k)$}
As seen in equation \ref{util} \& \ref{profit}, $p$ is a unit price, and the revenue for $j$th producer is equal to $p_j$ times $Q_j$ (the number of customers). $c$ is the fixed costs. Each of the parameters $c, \alpha, \beta, \gamma, k$ straightforwardly captures a marginal effect of the corresponding variable.

@@

\subsubsection{Trust update}
At each period, a consumer updates her trust levels in all the producers. Here, as well, we try to incorporate social embeddedness into the process. The idea is that the more embedded, the greater chance to interact with the producer and therefore increase the trust. As a result, we assume that an upward adjustment takes place with probability of $0.5 + \nicefrac{e_j}{2}$. Specifically, the $i$th consumer's trust in $j$th producer is updated as follows:
\begin{displaymath}
t'_{ij} =
\begin{cases}
t_{ij} \times 1.05 & \text{with probability } 0.5 + \dfrac{e_j}{2}\\
t_{ij} \times 0.99 & \text{otherwise}
\end{cases}
\end{displaymath}
The mild downward adjustment reflects gradual ``forgetting''. Finally, there is no trust update for conventional producers.

@@

Given this context at the beginning of every period each producer has a random chance to adjust their prices up or down depending on their profits from the previous period. The inclusion of a stochastic element is meant to model the infrequency of price adjustments due to either adjustment costs or simply neglect and the chances of an adjustment occurring are proportional to their situation. For those producers with negative profits the probability of adjustment is proportional to how far they missed their breakeven point (eg. a producer with costs of \$50 and revenue of \$20 has a $\nicefrac{30}{50}=60\%$ chance of adjustment). On the other hand producers with positive profits have chances proportional to how much they missed their ideal profit margin (eg. a producer with ideal profits of \$20 and with actual profits of \$12 has a $\nicefrac{8}{20}=40\%$ chance of adjustment). Notice that producers with profits above their ideal level will never raise their prices further, even if their sales are extremely high and they are functioning as a monopoly. This process tacitly acknowledges that producers in the local food sector tend to strive for a minimum level of profits as opposed to unfettered profit maximization.

As mentioned above, producers do not naively adjust their prices up or down, but instead take into account their environment just as a seller at a farmer's market stall responds to the prices of nearby vendors. If a producer finds itself with negative profits and decides to adjust their price downward they look in their local neighborhood and undercut the lowest price in an attempt to steal sales and boost their revenue. If they are already the lowest they keep their prices fixed (implicitly acknowledging that lower prices will only hurt them further). Positive profit producers act similarly; if they are adjusting prices they raise their price to undercut their neighbors assuming they find neighbors with higher prices. Otherwise they keep their prices fixed and continue.It is important to notice that price is the only choice variable for producers, and thus is their only means of responding to poor sales and negative profits. Therefore lone producers with negative profits will inevitably close, a perhaps unrealistic result. Future work may include additional response variables, such as adjusting quality or embeddedness, to better model a producer's ability to deal with losses.

\subsubsection{Entry \& Exit}
Since the number of producers is our metric of focus it is natural to include mechanisms for both entry and exit. The exit mechanism is simple and stark: at the beginning of each period any firm with a negative cash balance closes. In future models we may distinguish between shutdown and exit (urrently a meaningless distinction since our model lacks variable costs) but for now this mechanism will suffice.

>>>>

\subsubsection{Others $(p, c, \alpha, \beta, \gamma, k)$}
As seen in equation \ref{util} \& \ref{profit}, $p$ is a unit price, and the revenue for the $j$th producer is equal to $p_j$ times $Q_j$ (the total sales as well as the number of customers). $c$ is the fixed costs. Each of the parameters $c, \alpha, \beta, \gamma, k$ straightforwardly captures a marginal effect of the corresponding variable.

@@

\subsubsection{Trust update}
At each period, a consumer updates her trust levels in all the producers. Here, as well, we try to incorporate social embeddedness into the process. The idea is that the more embedded, the greater chance to interact with the producer and therefore increase the trust. As a result, we assume that an upward adjustment takes place with probability of $0.5 + \nicefrac{e_j}{2}$. Specifically, the $i$th consumer's trust in $j$th producer is updated as follows:
\begin{displaymath}
t'_{ij} =
\begin{cases}
t_{ij} \times 1.05 & \text{with probability } 0.5 + \dfrac{e_j}{2}\\
t_{ij} \times 0.99 & \text{otherwise}
\end{cases}
\end{displaymath}
The mild downward adjustment reflects gradual ``forgetting''. We implicitly set it unbounded from above and equal to 1 from below. The lower bound provides a limit of forgetting and an initial value for new producers. We are aware that both bounds should make sense relative to the other variables in equation \ref{util}. Finally, there is no trust update for conventional producers.

@@

Given this context, at the beginning of every period, each producer has a random chance to adjust their prices up or down depending on their profits from the previous period. The inclusion of a stochastic element is meant to model the infrequency of price adjustments due to either adjustment costs or simply neglect. The chances of an adjustment occurring are proportional to their situations. For those producers with negative profits, the probability of adjustment is proportional to how far they missed their breakeven point (e.g. a producer with costs of \$50 and revenue of \$20 has a $\nicefrac{30}{50}=60\%$ chance of adjustment). On the other hand, producers with positive profits have chances proportional to how much they missed their ideal profit margin (e.g. a producer with ideal profits of \$20 and with actual profits of \$12 has a $\nicefrac{8}{20}=40\%$ chance of adjustment). Notice that producers with profits above their ideal level will never raise their prices further, even if their sales are extremely high and they are functioning as a monopoly. This process tacitly acknowledges that producers in local food systems tend to strive for a minimum level of profits as opposed to unfettered profit maximization.

As mentioned above, producers do not naively adjust their prices up or down, but instead take into account their environment just as a seller at a farmers' market stall responds to the prices of nearby vendors. If producers find themselves with negative profits and decide to adjust their prices downward, they look in their local neighborhood and undercut the lowest price in an attempt to steal sales and boost their revenue. If they are already the lowest, they keep the prices fixed (implicitly acknowledging that lower prices will only hurt them further). Positive profit producers act similarly; if they are adjusting prices, they raise their prices to undercut their neighbors assuming they find neighbors with higher prices. Otherwise they keep their prices fixed and continue. It is important to notice that price is the only choice variable for producers, and thus is their only means of responding to poor sales and negative profits. Therefore lone producers with negative profits will inevitably close, perhaps an unrealistic result. Future work may include additional response variables, such as adjusting quality or embeddedness, to better model a producer's ability to deal with losses.

\subsubsection{Entry \& Exit}
Since the number of producers is our metric of focus it is natural to include mechanisms for both entry and exit. The exit mechanism is simple and stark: at the beginning of each period any firm with a negative cash balance closes. In future models we may distinguish between shutdown and exit (currently a meaningless distinction since our model lacks variable costs) but for now this mechanism will suffice.
